---
title: "Homework1"
author: "Laha Ale"
date: "1/27/2019"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,warning = FALSE)
```

## Exercise 7
#### a)

$P1 = (\theta_1,\lambda_1),P2 = (\theta_2,\lambda_2)$

$D=R \phi=Rarccos[sin\theta_1sin\theta_2+cos\theta_1cos\theta_2 cos(\lambda_1-\lambda_2)]$ \ \
```{r q7_a}
#R code block
geodesic <- function(P1,P2,R=6371)
{
  P1 <- P1*pi/180
  P2 <- P2*pi/180
  theta_1 <- P1[2]
  lambda_1 <- P1[1]
  theta_2 <- P2[2]
  lambda_2 <- P2[1]
  D <- R*acos(sin(theta_1)*sin(theta_2)+cos(theta_1)*cos(theta_2)*cos(lambda_1-lambda_2))
  return(D)
}
```
#### b)
```{r q7_b_1}
#R code block
Chicago <- c(87.63,41.88)
Minneapolis <- c(93.22,44.89)
geo_D_CM <- geodesic(Chicago,Minneapolis)
print(paste("The distance between Chicago and Minneapolis is",geo_D_CM))
```

```{r q7_b_2}
#R code block
New_York <- c(73.97,40.78)
New_Orleans <- c(90.25,29.98)
geo_D_NN <- geodesic(New_York,New_Orleans)
print(paste("The distance between New York and New Orleans is",geo_D_NN))
```

## Exercise 8


Suppose, $P1 = (\theta_1,\lambda_1),P2 = (\theta_2,\lambda_2)$\
Naive Euclidean Distance:\
$||P_1-P_2||\pi R/180=\pi R/180\sqrt{(\theta_2-\theta_1)^2+(\lambda_2-\lambda_1)^2}$ \ \
```{r q8_00}
#R code block
Naive_Euclidean<-function(P1,P2,R=6371)
{
  theta_1 <- P1[2]
  lambda_1 <- P1[1]
  theta_2 <- P2[2]
  lambda_2 <- P2[1]
  D <- sqrt((theta_2- theta_1)^2+(lambda_2-lambda_1)^2)*pi*R/180
  return(D)
}
```
```{r q8_11}
#R code block
NE_D_CM<- Naive_Euclidean(Chicago,Minneapolis)
print(paste("The Naive Euclidean distance between Chicago and Minneapolis is",NE_D_CM))
```
```{r q8_21}
#R code block
NE_D_NN <- Naive_Euclidean(New_York,New_Orleans)
print(paste("The Naive Euclidean distance between New York and New Orleans is",NE_D_NN))
```
As we can see from above, Naive Euclidean distance larger than geodesic distance.


## Exercise 9
Chordal distnace:\
$chordal=\sqrt{(x_2-x_1)^2+(y_2-y_1)^2+(z_2-z_1)^2}$ \ \
```{r q9_0}
#R code block
Chordal_D<-function(P1,P2,R=6371)
{
  P1 <- P1*pi/180
  P2 <- P2*pi/180
  theta_1 <- P1[2]
  lambda_1 <- P1[1]
  theta_2 <- P2[2]
  lambda_2 <- P2[1]
  x_1 <- R*cos(theta_1) * cos(lambda_1)
  y_1 <- R*cos(theta_1)* sin(lambda_1)
  z_1 <- R*sin(theta_1)
  x_2 <- R*cos(theta_2) * cos(lambda_2)
  y_2 <- R*cos(theta_2)* sin(lambda_2)
  z_2 <- R*sin(theta_2)
  D <- sqrt((x_2-x_1)^2+(y_2-y_1)^2+(z_2-z_1)^2)
  return(D)
}
```
```{r q9_1}
#R code block
C_D_CM <- Chordal_D(Chicago,Minneapolis)
print(paste("The Chordal distance between Chicago and Minneapolis is",C_D_CM))
C_D_NN <- Chordal_D(New_York,New_Orleans)
print(paste("The Chordal distance between New York and New Orleans is",C_D_NN))
```

\begin{center}
 \begin{tabular}{||c c c c||} 
 \hline
 Cities & Geodesic & Naive Euclidean & Chordal \\ [0.5ex] 
 \hline
 Chicago,Minneapolis & 561.99 & 705.96 & 561.81 \\ 
 \hline
 New York,New Orleans & 1897.21 & 2172.37 & 1890.21 \\
 \hline
\end{tabular}
\end{center}

As we can see from above, Naive Euclidean distance larger than geodesic distance, and geodesic distance than Chordal distance.


## Exercise 11

```{r q11_0, echo=FALSE}
Sys.setenv(GOOGLE_MAPS_API_KEY = "your google map key")
```

```{r q11_a}
#R code block
#Sys.setenv(GOOGLE_MAPS_API_KEY = "may google api key")
library(RgoogleMaps)
library(sp)
library(maptools)
library(rgdal)
scallops_url="https://www.counterpointstat.com/uploads/1/1/9/3/119383887/scallops.txt"
data <- read.table(scallops_url,header=TRUE)
hybrid <- GetMap.bbox(lonR = range(data$long), latR = range(data$lat),
                     size=c(640,640), maptype = "hybrid")
PlotOnStaticMap(hybrid)


SP_longlat <- SpatialPoints(coords = cbind(data$long,data$lat),
                            proj4string = CRS("+proj=longlat +ellps=WGS84"))
SP_utm <- spTransform(SP_longlat, CRS("+proj=utm +zone=13 +datum=WGS84"))
plot(SP_utm)

mobile <- GetMap.bbox(lonR = range(data$long), latR = range(data$lat),
                     size=c(640,640), maptype = "mobile")
PlotOnStaticMap(mobile)
terrain <- GetMap.bbox(lonR = range(data$long), latR = range(data$lat),
                     size=c(640,640), maptype = "terrain")
PlotOnStaticMap(terrain)
satellite <- GetMap.bbox(lonR = range(data$long), latR = range(data$lat),
                     size=c(640,640), maptype = "satellite")
PlotOnStaticMap(satellite)


```

