---
title: "Test 1 MATH 6344 SP 2019"
author: "Laha Ale"
date: "March 1, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### *Note: This file is produced by RMarkdown , and the lines start with ## are the outputs of R codes.*

##Prepare Data


```{r data}
library(spBayes)
library(classInt)
library(RColorBrewer)
library(geoR)
#Pollution ppm 
ppm <- c(14.3, 10.05,9.91,
         10.15,12.31,10.39,
         9.61,11.10,9.16,
         9.39,12.05)

#X Coordinate
X <- c(0.35,0.80,0.41,
       0.05,0.21,0.04,
       0.55,0.11,0.49,
       0.15,0.38)
#Y Coordinate
Y <- c(0.85,0.44,0.79,
       0.08,0.58,0.92,
       0.32,0.85,0.13,
       0.08,0.83)
# Construct dataframe
ppm.df <- data.frame("ppm"=ppm,"X"=X,"Y"=Y)
ppm.df
coords <- as.matrix(ppm.df[,c("X","Y")]) 
```

## 1) Show a 3d graph for these data and contour plot

### method 1 
```{r contour}
library(akima)
int.ppm <-interp.new(ppm.df$X,ppm.df$Y,ppm.df$ppm,extrap=TRUE)
image(int.ppm)
contour(int.ppm, add=T)
persp(int.ppm)
persp(int.ppm,theta = -70, phi = 45, d=1)
persp(int.ppm,theta = -70, phi = 45, d=1,col="red",shade=0.75,border=NA)
```

### method 2

```{r contour1}
library(MBA)
library(fields) ## For using the image.plot function
x.res <- 100
y.res <- 100
col.br <- colorRampPalette(c("blue", "cyan", "yellow", "red"))
col.pal <- col.br(5)


surf <- mba.surf(cbind(coords, ppm),
                 no.X = x.res, no.Y = y.res, h = 5,
                 m = 2, extend = FALSE)$xyz.est
image.plot(surf, xaxs = "r", yaxs = "r",
           xlab = "X", ylab = "Y", col = col.br(25))
contour(surf, add=T)
```
```{r contour2,eval=F}

library(rgl)
col <- rbind(0, cbind(matrix(drape.color(surf[[3]],
                                         col = col.br(25)), 
                                  x.res - 1, y.res - 1), 0))
surface3d(surf[[1]], surf[[2]], surf[[3]], col = col)
axes3d()

title3d(main = "PPM", xlab = "X", ylab = "Y", zlab = "ppm")
drape.plot(surf[[1]], surf[[2]], surf[[3]], col = col.br(150),
           theta = 225, phi = 50, border = FALSE, add.legend = FALSE, 
           xlab = "X", ylab = "Y", zlab = "ppm")
image.plot(zlim = range(surf[[3]], na.rm = TRUE), 
           legend.only = TRUE, horizontal = FALSE)
```
```{r 3d, echo=F}
library(OpenImageR)
image = readImage("3d.png")
imageShow(image)
```
Although the second method is fancier than the previous one,  it would take a long time to plot the 3D plot.

## 2) Fit a variogram model of your choice, use any kind of estimation(MLE, OLS, MOM) or if the results still do not agree with the empirical variogram you can also use eyefit. 

```{r variogram}
## Load some libraries we will use

bins = 20
max.dist <- max(iDist(coords))
ppm.vario <- variog(coords = coords, data = ppm, 
                       uvec = (seq(0, max.dist, length = bins)))
plot(ppm.vario)
```
As we can see from above the data is not really organized and it hardly use MLE, OLS, MOM to fit. Therefore, let's try to useing eyefit. 
```{r eyefit,eval=F}
eyefit(ppm.vario,silent=TRUE)
```
```{r ex6_3dd, echo=F}
library(OpenImageR)
image = readImage("eye_fit.png")
imageShow(image)
```
We will choose data given by above eyefit. Sill $\sigma^2=15.51$ , Range $\Phi=1.33$ and Nugget $\tau^2=0.47$. I know those parameters and models are not reasonably fit the data. The data seems not suitable for spatial statistics analysis.

```{r fit_vorgam, echo=F}
fit.ppm<- variofit(ppm.vario,cov.model="gaussian",
                       fix.nugget=FALSE, ini.cov.pars=c(15.51,1.33), 
                       nugget=0.47)
fit.ppm
```
