---
title: "Homework3"
author: "Laha Ale"
date: "February 20, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = F,warning = F)
```

### *Note: This file is produced by RMarkdown , and the lines start with ## are the outputs of R codes.*

```{r impprt_libs}
library(geoR)
library(spBayes)
```

# Excerise 6
## step 1:load data and plot variogram

```{r 6x_vario}
url <- "https://www.counterpointstat.com/uploads/1/1/9/3/119383887/myscallops.txt" 
myscallops <- read.table(url,header = T)
coords <- as.matrix(myscallops[,c("lat","long")]) 
lgcatch<- myscallops$lgcatch

bins = 50
max.dist <- 0.7*max(iDist(coords))
myscps.vario <- variog(coords = coords, data = lgcatch, 
                       uvec = (seq(0, max.dist, length = bins)))
plot(myscps.vario)
```

### step 2: adjust paramters with eyefit
```{r 6x_eyefit,eval=F}
eyefit(myscps.vario,silent=TRUE)
```
```{r ex6_3dd, echo=F}
library(OpenImageR)
image = readImage("6x_eyefit.png")
imageShow(image)
```

### step 3: computing varfit
According step 2 results, the following parameters should setting as $\sigma^2=5.34,\Phi=0.88$ and nugget=1.67. And $exponential$ model seems works fine.
```{r 6x_variofit}
fit.lgcatch<- variofit(myscps.vario,cov.model="exponential",
                       fix.nugget=FALSE, ini.cov.pars=c(5.34,0.88), 
                       nugget=1.67)
fit.lgcatch
```
### step 4: Kriging
According step 3 results, the following parameters should setting as $\sigma^2=4.6326,\Phi=0.1767$ and nugget=0.3802.
Although the value $\Phi=0.1767$ is not making sense in the above results if we compare to results from step 2, we will use this value to compute the last point of the lgcatch data.
```{r 6x_Kriging}
point<-krige.conv(coords = coords, data = lgcatch,loc=c(length(lgcatch),1),
                  krige=krige.control(cov.pars=c(4.6326,0.1767),
                                      cov.model="exponential",
                                      nugget=0.3802))
point
pred_low <-point$predict - 2*sqrt(point$krige.var)
pred_high <-point$predict + 2*sqrt(point$krige.var)
print(paste("The PI is between",pred_low,"and",pred_high))
```
### step 5: Summary Results
As we can see from above, the predict and varirance are 2.64 and 5.26; therefore, the PI with $95\%$ confident is beteewn $2.64-2\times\sqrt{5.26}= -1.94$ and $2.64+2\times\sqrt{5.26}=7.23$, more accurate numbers have been printed above.

# Excerise 7

### step 1:load data and plot variogram

```{r 7x_vario}
url_coal <- "https://www.counterpointstat.com/uploads/1/1/9/3/119383887/coal.ash.txt" 
coalash <- read.table(url_coal,header = T)


coords_coal <- as.matrix(coalash[,c("x","y")]) 
coal <- coalash$coal
vario.coal <- variog(coords = coords_coal, data = coal, uvec = (seq(0, length = bins)))

plot(vario.coal)
```

### step 2: adjust paramters with eyefit
```{r 7_eyefit,eval=F}
eyefit(vario.coal ,silent=TRUE)
```
```{r ex7_3dd, echo=F}
library(OpenImageR)
image = readImage("7x_eyefit.png")
imageShow(image)
```

### step 3: computing varfit
According step 2 results, the following parameters should setting as $\sigma^2=1.29,\Phi=15.57$ and nugget=0.92. However, the value shown in eyefit results is a little strange, it may more make sense if we choose about $\frac{1}{17}$ base on variogram.  In the following code, let's try both values of $\Phi$ and select the one produces more accurate results in predicting $\tau^2$ and $\sigma^2$.

And $exponential$ model seems works fine.
```{r 7ex_variofit}
fit.coal<- variofit(vario.coal,cov.model="exponential",
                    fix.nugget=FALSE,
                    ini.cov.pars=c(1.29,1/17), nugget=0.92)
fit.coal
fit.coal<- variofit(vario.coal,cov.model="exponential",
                    fix.nugget=FALSE,
                    ini.cov.pars=c(1.29,15.57), nugget=0.92)
fit.coal
```

### step 4: Kriging Prediction
According step 3 results, Even though set $\phi=\frac{1}{17}$ is make more sense but the model can produce more accurate values of $\tau^2$ and $\sigma^2$ by setting $\phi=15.57$. Therefore, we will adopt the results from the second option. And the following parameters should setting as $\sigma^2=1.2026,\Phi=11.1435$ and $\tau^2=1.0372$.

```{r 7x_Kriging}
point<-krige.conv(coords = coords_coal, data = coal,loc=c(length(coal),1),
                  krige=krige.control(cov.pars=c(1.2026,11.1435),cov.model="exponential",nugget=1.0372))
point
pred_low <-point$predict - 2*sqrt(point$krige.var)
pred_high <-point$predict + 2*sqrt(point$krige.var)
print(paste("The PI is between",pred_low,"and",pred_high))
```
### step 5: Summary Results
As we can see from above, the predict and varirance are 9.66 and 2.76; therefore, the PI with $95\%$ confident is beteewn $9.66-2\times\sqrt{2.76}= 6.34$ and $9.66+2\times\sqrt{2.76}=12.98$, more accurate numbers have been printed above.

# Using Matrix Method
# Exercise 6
Given same previous work and choose paramters as below:

```{r 6x_mtx}
# assign the coords to given_coords except last row

given_coords <- coords[1:dim(coords)[1]-1,]
myscallops.covmat<-varcov.spatial(coords = given_coords,
                                  cov.model = "exponential", 
                                 nugget = 1.67,
                                 cov.pars = c(5.34,0.88))
# just print myscallops.covmat will be too long!
# myscallops.covmat
# setup covariance matrix with point for prediction
gamma_all<-varcov.spatial(coords = coords,
                           cov.model = "exponential", 
                           nugget = 1.67,
                           cov.pars = c(5.34,0.88))
# we are interested in last column
gamma <- gamma_all$varcov[,ncol(gamma_all$varcov)]
cov <- gamma[length(gamma)]
gamma <- gamma[-length(gamma)]

z <- lgcatch[-length(lgcatch)]
mu <- mean(z)
m <- rep(mu, length(z))
m <- matrix(m,nrow=length(z),ncol=1)
z <- matrix(z,nrow=length(z),ncol=1)


y_pred<- mu +t(gamma)%*%solve(myscallops.covmat$varcov)%*%(z-m)
y_pred

# now compute variance

cov
var_pred<-cov -t(gamma)%*%solve(myscallops.covmat$varcov)%*%gamma
var_pred

pred_low <- y_pred - 2*sqrt(var_pred)
pred_high <- y_pred + 2*sqrt(var_pred)
print(paste("The PI is between",pred_low,"and",pred_high))
```
As we can see, from above results, the matrix method results is almost the same as the API function results.


# Exercise 7
Given same previous work and choose paramters as below:

```{r 7x_mtx}
# assign the coords to given_coords except last row

given_coords <- coords_coal[1:dim(coords_coal)[1]-1,]
coal.covmat<-varcov.spatial(coords = given_coords,
                                  cov.model = "exponential", 
                                 nugget = 1.0372,
                                 cov.pars = c(1.2026,11.1435))
# just print myscallops.covmat will be too long!
# myscallops.covmat
# setup covariance matrix with point for prediction
gamma_all<-varcov.spatial(coords = coords_coal,
                           cov.model = "exponential", 
                           nugget = 1.0372,
                           cov.pars = c(1.2026,11.1435))
# we are interested in last column
gamma <- gamma_all$varcov[,ncol(gamma_all$varcov)]
cov <- gamma[length(gamma)]
gamma <- gamma[-length(gamma)]

z <- coal[-length(coal)]
mu <- mean(z)
m <- rep(mu, length(z))
m <- matrix(m,nrow=length(z),ncol=1)
z <- matrix(z,nrow=length(z),ncol=1)


y_pred<- mu +t(gamma)%*%solve(coal.covmat$varcov)%*%(z-m)
y_pred

# now compute variance

cov
var_pred<-cov -t(gamma)%*%solve(coal.covmat$varcov)%*%gamma
var_pred

pred_low <- y_pred - 2*sqrt(var_pred)
pred_high <- y_pred + 2*sqrt(var_pred)
print(paste("The PI is between",pred_low,"and",pred_high))
```
As we can see, from above results, the matrix method results is clost to  the API function results to some extend.
