---
title: "Homework4"
author: "Laha Ale"
date: "March 17, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r ex11_lib}
library(spdep)
library(maps)
library(maptools)
library(classInt)
library(RColorBrewer)
```

## Excercise 11

### (a) 

```{r ex11_1}
state.sat.scores = read.table("state-sat.dat", header=F)
colnames(state.sat.scores) <- c("STATE","VERBAL","MATH","ELIGIBLE")
head(state.sat.scores)
```
```{r ex11_2}
# create listw
usa.state <- map(database="state", fill=TRUE, plot=FALSE)
state.ID <- sapply(strsplit(usa.state$names, ":"),
                   function(x) x[1])

usa.poly <- map2SpatialPolygons(usa.state, 
                               IDs=state.ID)
usa.nb <- poly2nb(usa.poly)
usa.listb <- nb2listw(usa.nb, style="B")
usa.listw <- nb2listw(usa.nb, style="W")

# train SAR model

x = ((state.sat.scores$STATE=="alaska") |
       (state.sat.scores$STATE=="hawaii") | 
       (state.sat.scores$STATE=="us"))

index = c(1:nrow(state.sat.scores))[x]
state.sat.scores = state.sat.scores[-index,]

# binnary weights
stat.sat.sar.b = spautolm(ELIGIBLE~ VERBAL,
                        data=state.sat.scores,
                        family="SAR", 
                        listw=usa.listb, 
                        zero.policy=TRUE)
summary(stat.sat.sar.b)
# row-normalized weights
stat.sat.sar.w = spautolm(ELIGIBLE~ VERBAL,
                        data=state.sat.scores,
                        family="SAR", 
                        listw=usa.listw, 
                        zero.policy=TRUE)
summary(stat.sat.sar.w)

```

As we can see from the above model that both VERBAL and MATH are significant since their P-Value is small enough; therefore, we shall keep both.
The binnary weights model is $ELIGIBLE= 390.60579-0.653476*VERBAL$ ,and The row-normalized weights model is $ELIGIBLE=387.62517-0.65812*VERBAL$. The ELIGIBLE has a negative relation with VERBAL. A possible reason VERBAL score has a negative relative with ELIGIBLE  is that the erea higher verbal score the higher the more competitive in terms of admission for the area.

### (b)
```{r ex11_b}
library(ggplot2)
library(ggmap)
library(mapdata)

states <- map_data("state")
states <- data.frame(states)

states[,"sat_range"] <- 0
names(states)[6] <- "sat"
state.sat.scores$STATE <- unique(states$region)
 
for(i in 1:nrow(states))
{
  for (j in 1:nrow(state.sat.scores))
  {
    if(grepl(state.sat.scores[j,]$STATE, states[i,]$region))
    {
      #sat <-  (state.sat.scores[j,]$VERBAL+state.sat.scores[j,]$MATH)/2
      sat <- state.sat.scores[j,]$VERBAL
      states[i,]$sat <- sat
      if(sat <= 503)
      {
        states[i,]$sat_range <- "<503"
      }else if(sat > 503 && sat <=525 )
      {
        states[i,]$sat_range <- "504-525"
      }else if(sat > 525 && sat <562 )
      {
        states[i,]$sat_range <- "526-562"
      }else
      {
        states[i,]$sat_range <- ">563"
      }
    }
  }
}

ggplot(data = states) +
  geom_polygon(aes(x = long,
                   y = lat, 
                   fill = sat_range,group=group),
                   color = "white") +
  coord_fixed(1.3)
```

