---
title: "Homework 6"
author: "Laha Ale"
date: "April 22, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = F,warning = F)
```

### *Note: This file is produced by RMarkdown , and the lines start with ## are the outputs of R codes.*

## Excercise 6

```{r lib}
#load libraries
library(spBayes)
library(MBA)
library(geoR)
library(fields)
library(sp)
library(maptools)
library(rgdal)
library(classInt)
library(lattice)
library(dplyr)
```

### a)

```{r ex_6a}
lithology_df <- read.csv("lithology.csv",header = T)
# drop the NA or miss data
lithology_df <- lithology_df %>% filter( !is.na(Thickness_ft) & 
                                           !is.na(Surf_Elevation_ft_amsl) &
                                           !is.na(A_B_Elevation_ft_amsl))
lithology_df$Surf_Elevation_ft_amsl <- as.numeric(lithology_df$Surf_Elevation_ft_amsl)
lithology_df$A_B_Elevation_ft_amsl <- as.numeric(lithology_df$A_B_Elevation_ft_amsl)

# Extract the coordinates
coords <- as.matrix(lithology_df[,c("Easting_ft","Northing_ft")])

x.res <- 200; y.res <- 200

surf <- mba.surf(cbind(coords, 
                       lithology_df$Thickness_ft), 
                 no.X=x.res, no.Y=y.res, h=5, 
                 m=2, extend=FALSE)$xyz.est
zlim.Thickness_ft <- range(surf[[3]], na.rm=TRUE)
image.plot(surf, xaxs = "r",
           yaxs = "r", xlab="Easting (ft)", 
           ylab="Northing (ft)",
           main="Thickness(ft)" )
points(coords)
contour(surf,add = T)

surf <- mba.surf(cbind(coords, 
                       lithology_df$Surf_Elevation_ft_amsl), 
                 no.X=x.res, no.Y=y.res, h=5, 
                 m=2, extend=FALSE)$xyz.est
zlim.Surf_Elevation_ft_amsl <- range(surf[[3]], na.rm=TRUE)
image.plot(surf, xaxs = "r",
           yaxs = "r", xlab="Easting (ft)", 
           ylab="Northing (ft)",
           main="Surf Elevation (ft amsl)")
points(coords)
contour(surf,add = T)


surf <- mba.surf(cbind(coords, 
                       lithology_df$A_B_Elevation_ft_amsl), 
                 no.X=x.res, no.Y=y.res, h=5, 
                 m=2, extend=FALSE)$xyz.est
zlim.A_B_Elevation_ft_amsl <- range(surf[[3]], na.rm=TRUE)
image.plot(surf, xaxs = "r",
           yaxs = "r", xlab="Easting (ft)", 
           ylab="Northing (ft)",
           main="A-B Elevation(ft amsl)")
points(coords)
contour(surf,add = T)
```

As we can see from the above results, the thickness IQR range about 5 feet to 20 feet whereas Surf Elevation is about 10 to 40 feet. Most of the points are located around Easting 265000 feet and Northing 67000 feet.

### b)

```{r ex_6b1}
# compute the variogram for exponential
log.thickness <- log(lithology_df$Thickness_ft)
bins = 50
max.dist <- 0.1*max(iDist(coords))
log.thick.vario <- variog(coords = coords, data = log.thickness, 
                       uvec = (seq(0, max.dist, length = bins)))
```
```{r 6x_eyefit,eval=F}
plot(log.thick.vario)
eyefit(log.thick.vario,silent=TRUE)
```
```{r ex6_3dd, echo=F}
library(OpenImageR)
image = readImage("exp_viro.png")
imageShow(image)
```
```{r ex_6b12}
# compute the variogram for matern
log.thick.vario <- variog(coords = coords, data = log.thickness, 
                       uvec = (seq(0, max.dist, length = bins)))
```
```{r 6x_eyefi2t,eval=F}
plot(log.thick.vario)
eyefit(log.thick.vario,silent=TRUE)
```
```{r ex6_3ddw, echo=F}
library(OpenImageR)
image = readImage("mattern_viro.png")
imageShow(image)
```

```{r ex_6b_exponential}

log.thickness <- log(lithology_df$Thickness_ft)

p <- 3 # This is the number of columns in the design matrix
# Set the prior mean and precision for the regression
beta.prior.mean <- as.matrix(rep(0, times=p))
beta.prior.precision <- matrix(0, nrow=p, ncol=p)

# For use with bayesGeostatExact, do the following
phi <- 0.014 ## Set the spatial range (from the variogram)
alpha <- 0.016/0.08 ## Set the nugget/partial-sill ratio
sigma.sq.prior.shape <- 0.1 ## Set IG shape for sigma.sq (partial sill)
sigma.sq.prior.rate <- 0.1 ## Set IG scale for sigma.sq (partial sill)

# Run bayesGeostatExact to deliver exact posterior samples
sp.exact <- bayesGeostatExact(
  log.thickness ~ Surf_Elevation_ft_amsl + A_B_Elevation_ft_amsl,
  data=lithology_df, coords=coords, n.samples=1000,
  beta.prior.mean=beta.prior.mean,
  beta.prior.precision=beta.prior.precision,
  cov.model="exponential",
  phi=phi, alpha=alpha,
  sigma.sq.prior.shape=sigma.sq.prior.shape,
  sigma.sq.prior.rate=sigma.sq.prior.rate,
  sp.effects=FALSE)
#Produce the posterior summaries
round(summary(sp.exact$p.samples)$quantiles,3)

```



```{r ex_6b_matern}

phi <- 1/1100 ## Set the spatial range (from the variogram)
alpha <- 0.094/0.58 ## Set the nugget/partial-sill ratio
nu <- 0.5

# Run bayesGeostatExact to deliver exact posterior samples
sp.exact <- bayesGeostatExact(
  log.thickness ~ Surf_Elevation_ft_amsl + A_B_Elevation_ft_amsl,
  data=lithology_df, coords=coords, n.samples=1000,
  beta.prior.mean=beta.prior.mean,
  beta.prior.precision=beta.prior.precision,
  cov.model="matern",
  phi=phi, alpha=alpha,
  nu=nu,
  sigma.sq.prior.shape=sigma.sq.prior.shape,
  sigma.sq.prior.rate=sigma.sq.prior.rate,
  sp.effects=FALSE)
round(summary(sp.exact$p.samples)$quantiles,3)
``` 
As we can see from the above results, the log Thickness has a negative correlation. For exponential, the model can draw as $log.thickness = 2.315-0.048*Surf\_Elevation\_ft\_amsl-0.015* A\_B\_Elevation\_ft\_amsl$, and $\sigma^2=0.276$, $\tau^2 = 0.055$. For exponential, the model can draw as $log.thickness = 1.926-0.039*Surf\_Elevation\_ft\_amsl -0.028* A\_B\_Elevation\_ft\_amsl$, and $\sigma^2=0.544$, $\tau^2 = 0.088$.

### c)
```{r ex_6c}

# Run spLM to deliver MCMC samples from marginal posterior distributions
n.samples <- 1000
thickness.sp <- spLM(log.thickness ~ Surf_Elevation_ft_amsl + A_B_Elevation_ft_amsl,
               data=lithology_df, coords=coords, 
               starting=list("phi"=3/1100,"sigma.sq"=0.08,"tau.sq"=0.02), 
               tuning=list("phi"=0.1, "sigma.sq"=0.05, "tau.sq"=0.05),
               priors=list("phi.Unif"=c(3/1500, 3/50),
                           "sigma.sq.IG"=c(0.1,0.1),"tau.sq.IG"=c(0.1, 0.1)), 
               cov.model="exponential",n.samples=n.samples)

round(summary(mcmc(thickness.sp$p.theta.samples))$quantiles,3)


# Recover spatial residuals using spRecover
burn.in <- floor(0.75*n.samples)
thickness.sp <- spRecover(thickness.sp, start=burn.in, thin=2)

# The posterior samples of the regression coefficients and the spatial effects can then be obtained as
beta.samples = thickness.sp$p.beta.recover.samples
w.samples = thickness.sp$p.w.recover.samples

# Obtain trace plots for regression coefficients
par(mfrow=c(2,2))
plot(beta.samples, auto.layout=TRUE, density=FALSE)
par(mfrow=c(1,1))
# Obtain posterior means and sd's of spatial residuals for each location
w.hat.mu <- apply(w.samples,1,mean)
w.hat.sd <- apply(w.samples,1,sd)


surf <- mba.surf(cbind(coords, w.hat.mu),
                 no.X=x.res, no.Y=y.res, extend=FALSE)$xyz.est
image.plot(surf, xaxs = "r", yaxs = "r", 
           main="Mean Spatial Effects")
contour(surf,add = T)

surf <- mba.surf(cbind(coords, w.hat.mu),
                 no.X=x.res, no.Y=y.res, extend=FALSE)$xyz.est
image.plot(surf, xaxs = "r", yaxs = "r", 
           zlim = zlim.Thickness_ft,
           main="log(Thickness) Mean Spatial Effects Over Thickness(ft)")
contour(surf,add = T)


surf <- mba.surf(cbind(coords, w.hat.mu),
                 no.X=x.res, no.Y=y.res, extend=FALSE)$xyz.est
image.plot(surf, xaxs = "r", yaxs = "r", 
           zlim = zlim.Surf_Elevation_ft_amsl,
           main="log(Thickness) Mean Spatial Effects Over Surf Elevation (ft amsl)")
contour(surf,add = T)
surf <- mba.surf(cbind(coords, w.hat.mu),
                 no.X=x.res, no.Y=y.res, extend=FALSE)$xyz.est
image.plot(surf, xaxs = "r", yaxs = "r", 
           zlim = zlim.A_B_Elevation_ft_amsl,
           main="log(Thickness) Mean Spatial Effects Over A-B Elevation(ft amsl) ")
contour(surf,add = T)
```

### d)

```{r ex_6d}
# run model with purely spatial
# note: set tau.sq=0 will work on spRecover and spDiag, therefore we set it a value close to 0
thickness.sp2 <- spLM(log.thickness ~ Surf_Elevation_ft_amsl + A_B_Elevation_ft_amsl,
               data=lithology_df, coords=coords, 
               starting=list("phi"=3/1100,"sigma.sq"=0.08,"tau.sq"=1E-10), 
               tuning=list("phi"=0.1, "sigma.sq"=0.05, "tau.sq"=1E-10),
               priors=list("phi.Unif"=c(3/1500, 3/50), 
                           "sigma.sq.IG"=c(0.1,0.1),
                           "tau.sq.IG"=c(1E-10, 1E-10)), 
               cov.model="exponential",n.samples=n.samples)

round(summary(mcmc(thickness.sp2$p.theta.samples))$quantiles,3)
thickness.sp2 <- spRecover(thickness.sp2, start=burn.in, thin=2)
Dic1 = spDiag(thickness.sp,start=burn.in,verbose=FALSE)
Dic1
Dic2 = spDiag(thickness.sp2,start=burn.in,verbose=FALSE)
Dic2
```
Compare the two models with above DIC values, the second model with much nugget is better than the first one with nugget.



## Excercise 7

### a)


```{r ex_7a}
bat_df <- read.table("batonrouge.dat",header = T)

# Extract the coordinates
coords <- as.matrix(bat_df[,c("Longitude","Latitude")])

# compute the variogram for exponential
bins = 50
max.dist <- 0.7*max(iDist(coords))
log.selling.vario <- variog(coords = coords, data = bat_df$logSellingPr, 
                          uvec = (seq(0, max.dist, length = bins)))
```
```{r ex7a_eyefit,eval=F}
plot(log.selling.vario)
eyefit(log.selling.vario,silent=TRUE)
```
```{r ex7_dd, echo=F}
library(OpenImageR)
image = readImage("exp_var_7a.png")
imageShow(image)
```

### b)
```{r ex_7b}
# OLS 
lm.log.selling = lm(logSellingPr~LivingArea + Age + OtherArea + Bathrooms, data=bat_df)
lm.log.selling
summary(lm.log.selling)
```
As we can see from the above results, the linear models is $logSellingPr = 10.2065918+0.0005892*LivingArea-0.0099027*Age+0.0001508*OtherArea-0.0512299*Bathrooms$, the model log selling price has a negative relationship with age and bathrooms. It is easy to understand that the old house should be cheaper, but it is a little strange about the bathrooms.  Mybe, the people, do not want to bathrooms occupy too much area of the house.


### c)


```{r ex_7c}
# Obtain OLS residuals
selling.resid = resid(lm.log.selling)
selling.resid.vario <- variog(coords = coords, data = selling.resid, 
                            uvec = (seq(0, max.dist, length = bins)))
```
```{r ex7c_eyefit,eval=F}
plot(selling.resid.vario)
eyefit(selling.resid.vario,silent=TRUE)
```
```{r ex7_dddc, echo=F}
library(OpenImageR)
image = readImage("exp_var_7c.png")
imageShow(image)
```

### d)


```{r ex_7d}
# the values of sigma square, tau sqaure  and nugget are from the above ressults
point<-krige.conv(coords = coords, data = bat_df$logSellingPr,loc=c(length(bat_df$logSellingPr),1),
                  krige=krige.control(cov.pars=c(0.22,0.17),cov.model="exponential",nugget=0.05))
point
pred_low <-point$predict - 2*sqrt(point$krige.var)
pred_high <-point$predict + 2*sqrt(point$krige.var)
print(paste("The 95% confident PI is between",pred_low,"and",pred_high))
```

### e)
Give above model:
$logSellingPr = 10.2065918+0.0005892*LivingArea-0.0099027*Age+0.0001508*OtherArea-0.0512299*Bathrooms$

```{r ex_7e}
LivingArea <- 938
Age <- 25
OtherArea <- 332
Bathrooms <- 3
logSellingPr <- 10.2065918+0.0005892*LivingArea-0.0099027*Age+0.0001508*OtherArea-0.0512299*Bathrooms
print(paste("predict results=",logSellingPr))
```
As we can see, the predict results is 10.40807, and the true value is 10.448, which is very good prediction.

### f)

```{r ex_7f}

n.samples = 1000

log.selling.mcmc <- spLM(logSellingPr~LivingArea + Age + OtherArea + Bathrooms, 
                         data=bat_df,
                         coords=coords, 
                         starting=list("phi"=0.17,"sigma.sq"=0.22,
                                       "tau.sq"=0.05), 
                         tuning=list("phi"=0.1, "sigma.sq"=0.05, "tau.sq"=0.05),
                         priors=list("phi.Unif"=c(3/1500, 10),
                                     "sigma.sq.IG"=c(0.1, 0.1),
                                     "tau.sq.IG"=c(0.1, 0.1)), 
                         cov.model="exponential",n.samples=n.samples)


round(summary(mcmc(log.selling.mcmc$p.theta.samples))$quantiles,3)

burn.in <- floor(0.75*n.samples)
log.selling.mcmc <- spRecover(log.selling.mcmc, start=burn.in, thin=2)
Dic1 = spDiag(log.selling.mcmc,start=burn.in,verbose=FALSE)
Dic1
```
The DIC is considerably small, but I cannot compare with other models, because other model donot have DIC.



### g)

```{r ex_7g}

beta.samples = log.selling.mcmc$p.beta.recover.samples
w.samples = log.selling.mcmc$p.w.recover.samples
## Obtain posterior means and sd's of spatial residuals for each location
w.hat.mu <- apply(w.samples,1,mean)
w.hat.mu
w.hat.sd <- apply(w.samples,1,sd)
w.hat.sd
summary(w.hat.mu)
summary(w.hat.sd)
```



